\documentclass[a4paper,11pt]{book}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{enumitem}
\usepackage{mathtools}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage[english]{babel}

% The minipage prevents LaTeX from putting linebreaks in lists
\newenvironment{symbollist} {\begin{minipage}{\linewidth}
    \begin{description}[labelindent=1em,itemsep=0em,parsep=0em]}
    {\end{description}\end{minipage}}

\newenvironment{caselist}[1][Case \arabic*:]{\begin{minipage}{\linewidth}
    \begin{enumerate}[label=\textbf{\emph{#1}},itemsep=0em,parsep=0.5em,leftmargin=*]}
    {\end{enumerate}\end{minipage}}

\newenvironment{trylist}{\begin{caselist}[Try:]}{\end{caselist}}

\newtheorem*{theorem}{Theorem}
% https://tex.stackexchange.com/questions/37472/spacing-before-and-after-with-newtheoremstyle
\newtheoremstyle{nospace}
  {0} % Space above
  {0} % Space below
  {} % Body font
  {} % Indent amount
  {\bfseries} % Theorem head font
  {.} % Punctuation after theorem head
  {.5em} % Space after theorem head
  {} % Theorem head spec (can be left empty, meaning `normal')
\theoremstyle{nospace} \newtheorem*{definition}{Definition}
\theoremstyle{definition} \newtheorem*{question}{Question}
\newtheorem*{solution}{Solution}
\newtheorem*{remark}{Remark}
\newtheorem*{example}{Example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% First page of book which contains 'stuff' like: %
%  - Book title, subtitle                         %
%  - Book author name                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Book's title and subtitle
\title{\Huge \textbf{Study Guide}  \\ \huge ECS 20: Discrete Math for CS
% Author
\author{\textsc{Computer Science Tutoring Club} \\ \textsc{Fall 2017}
}
}


\begin{document}
\frontmatter
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Auto-generated table of contents, list of figures and list of tables %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents

\mainmatter

%%%%%%%%%%%
% Preface %
%%%%%%%%%%%
\chapter*{How to Use This Study Guide}
\setlength{\parskip}{1em}
This study guide is meant to help you review topics before the midterms and the
final. Please do not use this study guide as your main source of preparation
for exams.

This study guide goes over each and every topic in ECS 20. I have previously
taken ECS 20 with Professor Koehl, so I have a pretty good idea about exams and
the material covered. Hence this study guide is tailor made for students in
Professor Patrice Koehl's ECS 20. If you are not in Professor Koehl's ECS 20
and still wish to use this study guide, please go ahead and do so. Please be
forewarned that some topics might not covered in this study guide that other
professors might cover (for example: Graph Theory is not covered in this study
guide).

Each chapter in the study guide reviews important concepts and theorems.
Furthermore, I have gone ahead and solved some questions that have been asked
on previous midterms. Additionally, I have added a few questions for you to try
out on your own (\textbf{No solutions have been provided}).

I sincerely hope that this study guide helps you better prepare for the
midterms and the final! Good luck on this challenging course.

\noindent \textbf{Aakash Prabhu (Class of 2019)} \\
\noindent \textbf{President, \emph{Computer Science Tutoring Club}}

\setlength{\parskip}{0em}

\input{Symbols}
\input{Logic}
\input{Proofs}
\input{Sets}
\input{Functions}
\input{Numbers}

\end{document}
